---
title: "Labor Economics Final Paper"
output: 
  html_document:
    theme: paper
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(usmap)
library(magrittr)
library(stargazer)

```

```{r}
# 
# 
# pulse_23_weight <- read_csv("https://www.dropbox.com/s/ptzdritvz1gudgo/pulse2021_repwgt_puf_23.csv?dl=1")
# 
# 
# pulse_22_weight <- read_csv("https://www.dropbox.com/s/ewengcdzm80rtz8/pulse2021_repwgt_puf_22.csv?dl=1")
# 
# pulse_21_weight <- read_csv("https://www.dropbox.com/s/ewengcdzm80rtz8/pulse2021_repwgt_puf_21.csv?dl=1")
# 
# pulse_20_weight <- read_csv("https://www.dropbox.com/s/ewengcdzm80rtz8/pulse2021_repwgt_puf_20.csv?dl=1")
# 
# pulse_19_weight <- read_csv("https://www.dropbox.com/s/ewengcdzm80rtz8/pulse2021_repwgt_puf_19.csv?dl=1")
# 
# pulse_18_weight <- read_csv("https://www.dropbox.com/s/ewengcdzm80rtz8/pulse2021_repwgt_puf_18.csv?dl=1")
```

```{r}

pulse_24 <- read_csv("pulse2021_puf_24.csv")

pulse_23 <- read_csv("https://www.dropbox.com/s/pini5g22oi6nrv0/pulse2021_puf_23.csv?dl=1")

pulse_22 <- read_csv("https://www.dropbox.com/s/8b1yyfzbpjwpakc/pulse2021_puf_22.csv?dl=1")

pulse_21 <- read_csv("https://www.dropbox.com/s/jqudrgvqnuzj37j/pulse2020_puf_21.csv?dl=1")

pulse_20 <- read_csv("https://www.dropbox.com/s/jjzxltc020ihy9i/pulse2020_puf_20.csv?dl=1")

pulse_19 <- read_csv("https://www.dropbox.com/s/j5mnfudk4vgnwx4/pulse2020_puf_19.csv?dl=1")

pulse_18 <- read_csv("https://www.dropbox.com/s/4m8tcz8s7catgpr/pulse2020_puf_18.csv?dl=1")



```

# Week 24

```{r}


# filter dataset to only include people in between ages of 21 and 25

young_24 <- pulse_24 %>%
  filter(TBIRTH_YEAR %in% (1996:2000))






# find % of people who did not do any work in past week


pct_young_work_24 <-  young_24 %>% 
  group_by(ANYWORK) %>% 
  summarise(weight = sum(PWEIGHT))


pct_young_work_24$weight[3] / 
    (pct_young_work_24$weight[1] + pct_young_work_24$weight[2] +  
     pct_young_work_24$weight[3])


# 38.8% of young people did not work in the past week
# filter out non substantive variables


young_24 <- young_24 %>% 
  filter(ANYWORK != -99)


pct_young_work_ms <- young_24 %>% 
  filter(EST_ST == 28)




pct_young_work_ms <-  pct_young_work_ms %>% 
  group_by(ANYWORK) %>% 
  summarise(weight = sum(PWEIGHT))

```

```{r}


pct_young_not_working_function_24 <- function(state_function_yp){
  st_filter <- young_24 %>% 
    filter(EST_ST == state_function_yp)
  
  not_working <- st_filter %>% 
    group_by(ANYWORK) %>% 
    summarise(weight_work = sum(PWEIGHT))
  
  pct_calc_young <- not_working$weight_work[2]/
  (not_working$weight_work[1] + not_working$weight_work[2])

#  pcdf <- data.frame(state_function, pct_calc)
  return(pct_calc_young)
}

# pct_young_not_working_function(38)



```

```{r}



statevector_youngemp <- sort(unique(pulse_24$EST_ST))

statevector_youngemp_df <- as.data.frame(statevector_youngemp)

statevector_youngemp_df <- statevector_youngemp_df %>% 
  mutate(row_id = row_number())



young_employment_mapped <- map(statevector_youngemp, 
                                pct_young_not_working_function_24)





young_employment_binded <- do.call("rbind", young_employment_mapped)




young_employment_binded_df <- as.data.frame(young_employment_binded)



young_employment_binded_df <- young_employment_binded_df %>% 
  mutate(row_id = row_number())




final_young_employment <- left_join(young_employment_binded_df, 
                                    statevector_youngemp_df,
                                    by = c("row_id" = "row_id"))


final_young_employment <- final_young_employment %>% 
  mutate(pct_not_working = round(V1 *100, 2))

final_young_employment_24 <- final_young_employment

```

# Week 23

```{r}

young_23 <- pulse_23 %>%
  filter(TBIRTH_YEAR %in% (1996:2000))


# find % of people who did not do any work in past week


pct_young_work_23 <-  young_23 %>% 
  group_by(ANYWORK) %>% 
  summarise(weight = sum(PWEIGHT))


pct_young_work_23$weight[3] / 
    (pct_young_work_23$weight[1] + pct_young_work_23$weight[2] +  
     pct_young_work_23$weight[3])


# 38.8% of young people did not work in the past week
# filter out non substantive variables


young_23 <- young_23 %>% 
  filter(ANYWORK != -99)


pct_young_work_ms_23 <- young_23 %>% 
  filter(EST_ST == 28)




pct_young_work_ms_23 <-  pct_young_work_ms_23 %>% 
  group_by(ANYWORK) %>% 
  summarise(weight = sum(PWEIGHT))

pct_young_work_ms_23


```

```{r}

young_23 <- pulse_23 %>%
  filter(TBIRTH_YEAR %in% (1996:2000))


# find % of people who did not do any work in past week


pct_young_work_23 <-  young_23 %>% 
  group_by(ANYWORK) %>% 
  summarise(weight = sum(PWEIGHT))


pct_young_work_23$weight[3] / 
    (pct_young_work_23$weight[1] + pct_young_work_23$weight[2] +  
     pct_young_work_23$weight[3])


# 38.8% of young people did not work in the past week
# filter out non substantive variables


young_23 <- young_23 %>% 
  filter(ANYWORK != -99)


pct_young_work_ms_23 <- young_23 %>% 
  filter(EST_ST == 30)




pct_young_work_ms_23 <-  pct_young_work_ms_23 %>% 
  group_by(ANYWORK) %>% 
  summarise(weight = sum(PWEIGHT))

pct_young_work_ms_23


```

```{r}


pct_young_not_working_function_23 <- function(state_function_yp){
  st_filter <- young_23 %>% 
    filter(EST_ST == state_function_yp)
  
  not_working <- st_filter %>% 
    group_by(ANYWORK) %>% 
    summarise(weight_work = sum(PWEIGHT))
  
  pct_calc_young <- not_working$weight_work[2]/
  (not_working$weight_work[1] + not_working$weight_work[2])

#  pcdf <- data.frame(state_function, pct_calc)
  return(pct_calc_young)
}

# pct_young_not_working_function(38)

```

```{r}

statevector_youngemp_23 <- sort(unique(pulse_23$EST_ST))

statevector_youngemp_df_23 <- as.data.frame(statevector_youngemp_23)

statevector_youngemp_df_23 <- statevector_youngemp_df_23 %>% 
  mutate(row_id = row_number())



young_employment_mapped_23 <- map(statevector_youngemp_23, 
                                pct_young_not_working_function_23)





young_employment_binded_23 <- do.call("rbind", young_employment_mapped_23)




young_employment_binded_df_23 <- as.data.frame(young_employment_binded_23)



young_employment_binded_df_23 <- young_employment_binded_df_23 %>% 
  mutate(row_id = row_number())




final_young_employment_23 <- left_join(young_employment_binded_df_23, 
                                    statevector_youngemp_df_23,
                                    by = c("row_id" = "row_id"))


final_young_employment_23 <- final_young_employment_23 %>% 
  mutate(pct_not_working = round(V1 *100, 2))

final_young_employment_23

```

# Week 22

```{r}

young_22 <- pulse_22 %>%
  filter(TBIRTH_YEAR %in% (1996:2000))


# find % of people who did not do any work in past week


pct_young_work_22 <-  young_22 %>% 
  group_by(ANYWORK) %>% 
  summarise(weight = sum(PWEIGHT))


pct_young_work_22$weight[3] / 
    (pct_young_work_22$weight[1] + pct_young_work_22$weight[2] +  
     pct_young_work_22$weight[3])


# 38.8% of young people did not work in the past week
# filter out non substantive variables


young_22 <- young_22 %>% 
  filter(ANYWORK != -99)


pct_young_work_ms_22 <- young_22 %>% 
  filter(EST_ST == 30)




pct_young_work_ms_22 <-  pct_young_work_ms_22 %>% 
  group_by(ANYWORK) %>% 
  summarise(weight = sum(PWEIGHT))

pct_young_work_ms_22


```

```{r}



pct_young_not_working_function_22 <- function(state_function_yp){
  st_filter <- young_22 %>% 
    filter(EST_ST == state_function_yp)
  
  not_working <- st_filter %>% 
    group_by(ANYWORK) %>% 
    summarise(weight_work = sum(PWEIGHT))
  
  pct_calc_young <- not_working$weight_work[2]/
  (not_working$weight_work[1] + not_working$weight_work[2])

#  pcdf <- data.frame(state_function, pct_calc)
  return(pct_calc_young)
}

# pct_young_not_working_function(38)

```

```{r}

statevector_youngemp_22 <- sort(unique(pulse_22$EST_ST))

statevector_youngemp_df_22 <- as.data.frame(statevector_youngemp_22)

statevector_youngemp_df_22 <- statevector_youngemp_df_22 %>% 
  mutate(row_id = row_number())



young_employment_mapped_22 <- map(statevector_youngemp_22, 
                                pct_young_not_working_function_22)





young_employment_binded_22 <- do.call("rbind", young_employment_mapped_22)




young_employment_binded_df_22 <- as.data.frame(young_employment_binded_22)



young_employment_binded_df_22 <- young_employment_binded_df_22 %>% 
  mutate(row_id = row_number())




final_young_employment_22 <- left_join(young_employment_binded_df_22, 
                                    statevector_youngemp_df_22,
                                    by = c("row_id" = "row_id"))


final_young_employment_22 <- final_young_employment_22 %>% 
  mutate(pct_not_working = round(V1 *100, 2))

final_young_employment_22


```

# Week 21

```{r}


young_21 <- pulse_21 %>%
  filter(TBIRTH_YEAR %in% (1996:2000))


# find % of people who did not do any work in past week


pct_young_work_21 <-  young_21 %>% 
  group_by(ANYWORK) %>% 
  summarise(weight = sum(PWEIGHT))


pct_young_work_21$weight[3] / 
    (pct_young_work_21$weight[1] + pct_young_work_21$weight[2] +  
     pct_young_work_21$weight[3])


# 38.8% of young people did not work in the past week
# filter out non substantive variables


young_21 <- young_21 %>% 
  filter(ANYWORK != -99)


pct_young_work_ms_21 <- young_21 %>% 
  filter(EST_ST == 30)




pct_young_work_ms_21 <-  pct_young_work_ms_21 %>% 
  group_by(ANYWORK) %>% 
  summarise(weight = sum(PWEIGHT))

pct_young_work_ms_21

```

```{r}


pct_young_not_working_function_21 <- function(state_function_yp){
  st_filter <- young_21 %>% 
    filter(EST_ST == state_function_yp)
  
  not_working <- st_filter %>% 
    group_by(ANYWORK) %>% 
    summarise(weight_work = sum(PWEIGHT))
  
  pct_calc_young <- not_working$weight_work[2]/
  (not_working$weight_work[1] + not_working$weight_work[2])

#  pcdf <- data.frame(state_function, pct_calc)
  return(pct_calc_young)
}

# pct_young_not_working_function(38)

```

```{r}


statevector_youngemp_21 <- sort(unique(pulse_21$EST_ST))

statevector_youngemp_df_21 <- as.data.frame(statevector_youngemp_21)

statevector_youngemp_df_21 <- statevector_youngemp_df_21 %>% 
  mutate(row_id = row_number())



young_employment_mapped_21 <- map(statevector_youngemp_21, 
                                pct_young_not_working_function_21)





young_employment_binded_21 <- do.call("rbind", young_employment_mapped_21)




young_employment_binded_df_21 <- as.data.frame(young_employment_binded_21)



young_employment_binded_df_21 <- young_employment_binded_df_21 %>% 
  mutate(row_id = row_number())




final_young_employment_21 <- left_join(young_employment_binded_df_21, 
                                    statevector_youngemp_df_21,
                                    by = c("row_id" = "row_id"))


final_young_employment_21 <- final_young_employment_21 %>% 
  mutate(pct_not_working = round(V1 *100, 2))

final_young_employment_21



```

# Week 20

```{r}


young_20 <- pulse_20 %>%
  filter(TBIRTH_YEAR %in% (1996:2000))


# find % of people who did not do any work in past week


pct_young_work_20 <-  young_20 %>% 
  group_by(ANYWORK) %>% 
  summarise(weight = sum(PWEIGHT))


pct_young_work_20$weight[3] / 
    (pct_young_work_20$weight[1] + pct_young_work_20$weight[2] +  
     pct_young_work_20$weight[3])


# 38.8% of young people did not work in the past week
# filter out non substantive variables


young_20 <- young_20 %>% 
  filter(ANYWORK != -99)


pct_young_work_ms_20 <- young_20 %>% 
  filter(EST_ST == 30)




pct_young_work_ms_20 <-  pct_young_work_ms_20 %>% 
  group_by(ANYWORK) %>% 
  summarise(weight = sum(PWEIGHT))

pct_young_work_ms_20

```

```{r}


pct_young_not_working_function_20 <- function(state_function_yp){
  st_filter <- young_20 %>% 
    filter(EST_ST == state_function_yp)
  
  not_working <- st_filter %>% 
    group_by(ANYWORK) %>% 
    summarise(weight_work = sum(PWEIGHT))
  
  pct_calc_young <- not_working$weight_work[2]/
  (not_working$weight_work[1] + not_working$weight_work[2])

#  pcdf <- data.frame(state_function, pct_calc)
  return(pct_calc_young)
}

# pct_young_not_working_function(38)


```

```{r}


statevector_youngemp_20 <- sort(unique(pulse_20$EST_ST))

statevector_youngemp_df_20 <- as.data.frame(statevector_youngemp_20)

statevector_youngemp_df_20 <- statevector_youngemp_df_20 %>% 
  mutate(row_id = row_number())



young_employment_mapped_20 <- map(statevector_youngemp_20, 
                                pct_young_not_working_function_20)





young_employment_binded_20 <- do.call("rbind", young_employment_mapped_20)




young_employment_binded_df_20 <- as.data.frame(young_employment_binded_20)



young_employment_binded_df_20 <- young_employment_binded_df_20 %>% 
  mutate(row_id = row_number())




final_young_employment_20 <- left_join(young_employment_binded_df_20, 
                                    statevector_youngemp_df_20,
                                    by = c("row_id" = "row_id"))


final_young_employment_20 <- final_young_employment_20 %>% 
  mutate(pct_not_working = round(V1 *100, 2))

final_young_employment_20




```

# Week 19

```{r}

young_19 <- pulse_19 %>%
  filter(TBIRTH_YEAR %in% (1996:2000))


# find % of people who did not do any work in past week


pct_young_work_19 <-  young_19 %>% 
  group_by(ANYWORK) %>% 
  summarise(weight = sum(PWEIGHT))


pct_young_work_19$weight[3] / 
    (pct_young_work_19$weight[1] + pct_young_work_19$weight[2] +  
     pct_young_work_19$weight[3])


# 38.8% of young people did not work in the past week
# filter out non substantive variables


young_19 <- young_19 %>% 
  filter(ANYWORK != -99)


pct_young_work_ms_19 <- young_19 %>% 
  filter(EST_ST == 30)




pct_young_work_ms_19 <-  pct_young_work_ms_19 %>% 
  group_by(ANYWORK) %>% 
  summarise(weight = sum(PWEIGHT))

pct_young_work_ms_19


```

```{r}


pct_young_not_working_function_19 <- function(state_function_yp){
  st_filter <- young_19 %>% 
    filter(EST_ST == state_function_yp)
  
  not_working <- st_filter %>% 
    group_by(ANYWORK) %>% 
    summarise(weight_work = sum(PWEIGHT))
  
  pct_calc_young <- not_working$weight_work[2]/
  (not_working$weight_work[1] + not_working$weight_work[2])

#  pcdf <- data.frame(state_function, pct_calc)
  return(pct_calc_young)
}

# pct_young_not_working_function(38)


```

```{r}


statevector_youngemp_19 <- sort(unique(pulse_19$EST_ST))

statevector_youngemp_df_19 <- as.data.frame(statevector_youngemp_19)

statevector_youngemp_df_19 <- statevector_youngemp_df_19 %>% 
  mutate(row_id = row_number())



young_employment_mapped_19 <- map(statevector_youngemp_19, 
                                pct_young_not_working_function_19)





young_employment_binded_19 <- do.call("rbind", young_employment_mapped_19)




young_employment_binded_df_19 <- as.data.frame(young_employment_binded_19)



young_employment_binded_df_19 <- young_employment_binded_df_19 %>% 
  mutate(row_id = row_number())




final_young_employment_19 <- left_join(young_employment_binded_df_19, 
                                    statevector_youngemp_df_19,
                                    by = c("row_id" = "row_id"))


final_young_employment_19 <- final_young_employment_19 %>% 
  mutate(pct_not_working = round(V1 *100, 2))

final_young_employment_19





```

# Week 18

```{r}

young_18 <- pulse_18 %>%
  filter(TBIRTH_YEAR %in% (1996:2000))


# find % of people who did not do any work in past week


pct_young_work_18 <-  young_18 %>% 
  group_by(ANYWORK) %>% 
  summarise(weight = sum(PWEIGHT))


pct_young_work_18$weight[3] / 
    (pct_young_work_18$weight[1] + pct_young_work_18$weight[2] +  
     pct_young_work_18$weight[3])


# 38.8% of young people did not work in the past week
# filter out non substantive variables


young_18 <- young_18 %>% 
  filter(ANYWORK != -99)


pct_young_work_ms_18 <- young_18 %>% 
  filter(EST_ST == 30)




pct_young_work_ms_18 <-  pct_young_work_ms_18 %>% 
  group_by(ANYWORK) %>% 
  summarise(weight = sum(PWEIGHT))

pct_young_work_ms_18



```

```{r}

pct_young_not_working_function_18 <- function(state_function_yp){
  st_filter <- young_18 %>% 
    filter(EST_ST == state_function_yp)
  
  not_working <- st_filter %>% 
    group_by(ANYWORK) %>% 
    summarise(weight_work = sum(PWEIGHT))
  
  pct_calc_young <- not_working$weight_work[2]/
  (not_working$weight_work[1] + not_working$weight_work[2])

#  pcdf <- data.frame(state_function, pct_calc)
  return(pct_calc_young)
}

# pct_young_not_working_function(38)



```

```{r}

statevector_youngemp_18 <- sort(unique(pulse_18$EST_ST))

statevector_youngemp_df_18 <- as.data.frame(statevector_youngemp_18)

statevector_youngemp_df_18 <- statevector_youngemp_df_18 %>% 
  mutate(row_id = row_number())



young_employment_mapped_18 <- map(statevector_youngemp_18, 
                                pct_young_not_working_function_18)





young_employment_binded_18 <- do.call("rbind", young_employment_mapped_18)




young_employment_binded_df_18 <- as.data.frame(young_employment_binded_18)



young_employment_binded_df_18 <- young_employment_binded_df_18 %>% 
  mutate(row_id = row_number())




final_young_employment_18 <- left_join(young_employment_binded_df_18, 
                                    statevector_youngemp_df_18,
                                    by = c("row_id" = "row_id"))


final_young_employment_18 <- final_young_employment_18 %>% 
  mutate(pct_not_working = round(V1 *100, 2))

final_young_employment_18






```

# Now combine all final data frames

```{r}


final_young_employment_18$statevector_youngemp_18
final_young_employment_18$pct_not_working

final_young_employment_18 <- final_young_employment_18 %>% 
  rename(state_fips = statevector_youngemp_18,
          pct_not_working_18 = pct_not_working )


final_young_employment_19 <- final_young_employment_19 %>% 
  rename(state_fips = statevector_youngemp_19,
         pct_not_working_19 = pct_not_working )


final_young_employment_20 <- final_young_employment_20 %>% 
  rename(state_fips = statevector_youngemp_20,
         pct_not_working_20 = pct_not_working )


final_young_employment_21 <- final_young_employment_21 %>% 
  rename(state_fips = statevector_youngemp_21,
         pct_not_working_21 = pct_not_working )


final_young_employment_22 <- final_young_employment_22 %>% 
  rename(state_fips = statevector_youngemp_22,
         pct_not_working_22 = pct_not_working )


final_young_employment_23 <- final_young_employment_23 %>% 
  rename(state_fips = statevector_youngemp_23,
         pct_not_working_23 = pct_not_working )


final_young_employment_24 <- final_young_employment_24 %>% 
  rename(state_fips = statevector_youngemp,
         pct_not_working_24 = pct_not_working)




```

```{r}


merged_final_young_emp_dataset <- final_young_employment_18 %>%
    left_join(final_young_employment_19, by='state_fips') %>%
    left_join(final_young_employment_20, by='state_fips') %>% 
    left_join(final_young_employment_21, by='state_fips') %>% 
    left_join(final_young_employment_22, by='state_fips') %>% 
    left_join(final_young_employment_23, by='state_fips') %>% 
    left_join(final_young_employment_24, by='state_fips')

head(merged_final_young_emp_dataset)


merged_final_young_emp_dataset <- merged_final_young_emp_dataset %>% 
  select(-starts_with('V1'))

head(merged_final_young_emp_dataset)

```

Need to pivot this table

```{r}

piv_young_emp_dataset <- merged_final_young_emp_dataset %>% 
  pivot_longer(c(pct_not_working_18, pct_not_working_19, 
                 pct_not_working_20, pct_not_working_21, 
                 pct_not_working_22, pct_not_working_23,
                 pct_not_working_24), names_to = "Wave",
               values_to = "pct_young_nw")

head(piv_young_emp_dataset)

piv_young_emp_dataset$wave_num <- piv_young_emp_dataset$Wave %>% 
  str_remove("pct_not_working_")

head(piv_young_emp_dataset)

piv_young_emp_dataset <- piv_young_emp_dataset %>% 
  select(c(state_fips, row_id, Wave, pct_young_nw, wave_num))

head(piv_young_emp_dataset)
```

# Plot

```{r}

piv_young_emp_grouped <- piv_young_emp_dataset %>% 
  group_by(state_fips)

piv_young_emp_grouped$wave_num <- 
  as.numeric(piv_young_emp_grouped$wave_num)


ggplot(piv_young_emp_grouped, aes(x = wave_num, y = pct_young_nw, 
                                  color = state_fips)) +
  geom_line() + 
  theme_minimal() +
  labs(x = "Wave", y = "% Young Not Working") +
  facet_wrap(~state_fips)


```

# Now, get other state level data

start with employment rates leading up to covid

```{r}

state_unemp_data_hist <- read_csv("county_unemp_data.csv")


state_unemp_data_hist <- state_unemp_data_hist %>% 
  filter(substring(fips_txt, 5, 5) == 0) %>% 
  filter(is.na(Rural_urban_continuum_code_2013)) %>% 
  filter(is.na(Urban_influence_code_2013)) %>% 
  filter(is.na(Metro_2013)) %>% 
  filter(Stabr != 'US') 


  
state_unemp_data_hist <- state_unemp_data_hist %>% 
  select(c(fips_txt, Stabr, Unemployment_rate_2010,
           Unemployment_rate_2011, Unemployment_rate_2012, 
           Unemployment_rate_2013, Unemployment_rate_2014,
           Unemployment_rate_2015, Unemployment_rate_2016,
           Unemployment_rate_2017, Unemployment_rate_2018,
           Unemployment_rate_2019))

head(state_unemp_data_hist)


piv_state_unemp_data_hist <- state_unemp_data_hist %>% 
  pivot_longer(!c(fips_txt, Stabr), names_to = "year", 
               values_to = "unemp_rate")

head(piv_state_unemp_data_hist)

```

Just take mean of state unemployment rate-- don't really want it time-variant

```{r}

#piv_state_unemp_data_hist$

state_mean_unemp <- piv_state_unemp_data_hist %>% 
  group_by(Stabr) %>% 
  mutate(mean_unemp = mean(unemp_rate)) %>% 
  select(c(fips_txt, Stabr, mean_unemp))

state_mean_unemp <- unique(state_mean_unemp)

state_mean_unemp$fips_txt


state_mean_unemp$fips_txt <- str_sub(state_mean_unemp$fips_txt, end=-4)

state_mean_unemp
  

```

Join the data sets

```{r}

head(piv_young_emp_dataset)

full_dataset <- left_join(piv_young_emp_dataset, state_mean_unemp,
                          by = c("state_fips" = "fips_txt"))

head(full_dataset)

```

```{r}

edu_data <- read_csv("county_edu.csv")

names(edu_data)

edu_data$`FIPS Code`

edu_data <- edu_data %>% 
  select(`FIPS Code`, State, `Area name` ,
         `2003 Rural-urban Continuum Code`, 
         `2003 Urban Influence Code`, `2013 Rural-urban Continuum Code`,
         `Percent of adults with a bachelor's degree or higher, 2015-19`)

head(edu_data)

edu_data <- edu_data %>% 
  filter(substring(`FIPS Code`, 5, 5) == 0) %>% 
  filter(substring(`FIPS Code`, 4, 4) == 0) %>%
  filter(is.na(`2003 Rural-urban Continuum Code`)) %>% 
  filter(is.na(`2003 Urban Influence Code`)) %>%
  filter(State != 'US') %>% 
  filter(!is.na(`Percent of adults with a bachelor's degree or higher, 2015-19`))

head(edu_data)



edu_data$`FIPS Code` <- str_sub(edu_data$`FIPS Code`, end=-4)

edu_data


head(full_dataset)

full_dataset <- left_join(full_dataset, edu_data, 
                          by = c("state_fips" = "FIPS Code"))

names(full_dataset)

full_dataset <- full_dataset %>% 
  select(c(state_fips, Wave, pct_young_nw, wave_num, Stabr, 
           `Percent of adults with a bachelor's degree or higher, 2015-19`,
           mean_unemp))

head(full_dataset)

full_dataset <- full_dataset %>% 
  rename(pct_with_ba = `Percent of adults with a bachelor's degree or higher, 2015-19`)

head(full_dataset)

```

# First pass model

```{r}


dataset_for_modeling <- full_dataset %>% 
  group_by(state_fips) %>% 
  summarise(mean_pct_young_nw = mean(pct_young_nw))


dataset_for_modeling <- left_join(dataset_for_modeling, full_dataset,
                                  by = c("state_fips" = "state_fips"))

head(dataset_for_modeling)

dataset_for_modeling <- dataset_for_modeling %>% 
  select(c(state_fips, mean_pct_young_nw, Stabr, pct_with_ba, mean_unemp))

head(dataset_for_modeling)

dataset_for_modeling <- distinct(dataset_for_modeling)


head(dataset_for_modeling)

beg_model <- lm(pct_young_nw ~ mean_unemp + pct_with_ba, data = full_dataset)

summary(beg_model)

stargazer(beg_model, type = "html", out = "labor_econ_model.html",
          dep.var.labels = "Percent of Young People not Working",
          covariate.labels = c("Mean Unemployment Rate", 
                               "Percent of Adults Holding a BA"),
          style = "aer")


```

# Make a map of state averages for time trends, Don't have to worry about cluttered facet map

```{r}

wave_avgs_young_emp <- piv_young_emp_dataset %>% 
  group_by(Wave) %>% 
  summarise(wave_avg = mean(pct_young_nw))



wave_avgs_young_emp$wave_num <- wave_avgs_young_emp$Wave %>% 
  str_remove("pct_not_working_")



wave_avgs_young_emp$wave_num <- as.numeric(wave_avgs_young_emp$wave_num)


ggplot(wave_avgs_young_emp, aes(x = wave_num, y = wave_avg)) +
  labs(x = "Wave", y = "% of Young People Responding 'not employed'",
       title = "Average Level of Young Unemployment of All States",
       caption = "Source: US Census Household Pulse Survey") +
  expand_limits(y=0) +
  theme_bw() +
  geom_line()


```
